---
title: "Untitled"
output: html_document
date: "2023-06-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(afex)
library(BayesFactor)
library(pwr)
load("UNM02_proc_data.RData")
```

```{r}
test_data <-  mutate(test_data, 
               certainty = case_when(cue_type == "C_NP" | cue_type == "C_P" ~ "certain",
                                    cue_type == "U_NP" | cue_type == "U_P" ~ "uncertain"),
               predictiveness = case_when(cue_type == "U_P" | cue_type == "C_P" ~ "predictive",
                                    cue_type == "U_NP" | cue_type == "C_NP" ~ "nonpreditive"))
mem_mean_parti <- filter(test_data, acc == 1) %>%
  group_by (pNum, certainty, predictiveness) %>%
  summarise(mem_score = mean(mem_score, na.rm = TRUE))
#now factorize the IV
mem_mean_parti$pNum <- factor(mem_mean_parti$pNum)
mem_mean_parti$certainty <- factor(mem_mean_parti$certainty)
mem_mean_parti$predictiveness <- factor(mem_mean_parti$predictiveness)
#ANOVA one between subjects factor (cue_type) on DV mem_score
mem_score_test_ANOVA <- aov_car(formula = mem_score ~ Error(pNum/certainty*predictiveness), data = mem_mean_parti, anova_table = list(es = "pes"))
print(mem_score_test_ANOVA)
```
Certainty: 
Very low power (0.0789029). If we want to run it between subjects, with the partial eta square reported, a 0.8 power, 1 df in the numerator y 2 groups, the total sample size needed would be 1303. 

Predictiveness:
Even lower power (0.0545615).  If we want to run it between subjects, with the partial eta square reported, a 0.8 power, 1 df in the numerator y 2 groups, the total sample size needed would be 7843.

Interaction:
according to G*Power, the power is 0.9957627.
So, if we want to run it between subjects, with the partial eta square reported, a 0.8 power, 1 df in the numerator y 2 groups, the total sample size needed would be 25. 
This is a bit odd, because is less of what we have used. Not sure this is right.


